Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-09 latex/main.tex
Fix latex compile

\begin{figure}[h]
    \centering
    \includegraphics[width=1\linewidth]{fig_recombinant_evidence.pdf}
    \caption{Evidence for recombination in the ARG. All plots show ``robust'' sc2ts recombination nodes, colored according
to their classification by rebar (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labelled starting with `X' and
those identified by Jackson et al.\cite{Jackson2021} are prefixed with a `J', e.g.
Jackson Group B (JGB) and Jackson Singleton recombinant 1 (JS1). 
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, XBB, XBB.1, and the recombinant origin of XZ+XAC+XAD+XAE+XAP;
each recombination node is a mix of variants matching the left parent (green), the right parent (blue), and de-novo mutations (gold).}
    \label{fig:averted_mutations}
\end{figure}

@@

\subsubsection{Mutational spectra}
To validate the mutational information contained in the sc2ts ARG,
we compared the all-site mutational spectra of major VOCs in the ARG,
a state-of-the-art UShER phylogeny built using the Viridian dataset\cite{Hunt2024}, and
a global UShER phylogeny analyzed in a previous study\cite{Bloom2023}.
Mutational spectra are relative proportions of each type of nucleotide mutation.

@@

\subsubsection{Concordance among recombination detection methods}
For RecombinHunt, we obtained two sets of detection results. 
One set is based on mutational profiles calculated from a global GISAID dataset, 
and the other set from a global Nextstrain dataset. 
We adjusted the RecombinHunt-identified breakpoint interval 
so that they are right-exclusive by adding one to the right coordinate.

@@

% See https://github.com/jeromekelleher/sc2ts-paper/issues/472
\begin{table}
\caption{Major deletions in Li et al. (2023).}
\label{tab:indels}
\begin{tabular}{rrcrrrr}
\toprule & \multicolumn{4}{c}{Li et al. (2023)} & \multicolumn{2}{c}{Sc2ts ARG} \\
\cmidrule(lr){1-5} \cmidrule(lr){6-7}
Start & Length & Region & Count & Frequency (%) & Count & Frequency (%)\\
\midrule
28362 & 9 & N & 3939164 & 43.0525 & 871774 & 35.1216 \\
28271 & 1 & Intergenic & 3853692 & 42.1183 & 1118363 & 45.0561 \\
21765 & 6 & S & 3179887 & 34.7541 & 339681 & 13.6849 \\
11288 & 9 & ORF1ab-nsp6 & 3040576 & 33.2315 & 290408 & 11.6998 \\
28248 & 6 & ORF8 & 3026186 & 33.0742 & 1114669 & 44.9073 \\
22029 & 6 & S & 3010665 & 32.9046 & 1112410 & 44.8163 \\
11283 & 9 & ORF1ab-nsp6 & 2011934 & 21.9891 & 40 & 0.0016 \\
6513 & 3 & ORF1ab-nsp3 & 1984683 & 21.6913 & 340765 & 13.7286 \\
21633 & 9 & S & 1941962 & 21.2244 & 532072 & 21.4359 \\
21987 & 9 & S & 1918094 & 20.9635 & N/A & N/A \\
22194 & 3 & S & 1833050 & 20.0340 & 312706 & 12.5982 \\
21991 & 3 & S & 912944 & 9.9779 & 288831 & 11.6363 \\
686 & 9 & ORF1ab-nsp1 & 115450 & 1.2618 & N/A & N/A \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=1\linewidth]{recombinant_evidence.pdf}
    \caption{Evidence for recombination in the ARG. All plots show ``robust'' sc2ts recombination nodes, colored according
to their classification by rebar (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labelled starting with `X' and
those identified by Jackson et al.\cite{Jackson2021} are prefixed with a `J', e.g.
Jackson Group B (JGB) and Jackson Singleton recombinant 1 (JS1). 
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, XBB, XBB.1, and the recombinant origin of XZ+XAC+XAD+XAE+XAP;
each recombination node is a mix of variants matching the left parent (green), the right parent (blue), and de-novo mutations (gold).}
    \label{fig:averted_mutations}
\end{figure}

@@

\subsubsection*{Mutational spectra}
To validate the mutational information contained in the sc2ts ARG,
we compared the all-site mutational spectra of major VOCs in the ARG,
a state-of-the-art UShER phylogeny built using the Viridian dataset\cite{Hunt2024}, and
a global UShER phylogeny analyzed in a previous study\cite{Bloom2023}.
Mutational spectra are relative proportions of each type of nucleotide mutation.

@@

\subsubsection*{Concordance among recombination detection methods}
For RecombinHunt, we obtained two sets of detection results. 
One set is based on mutational profiles calculated from a global GISAID dataset, 
and the other set from a global Nextstrain dataset. 
We adjusted the RecombinHunt-identified breakpoint interval 
so that they are right-exclusive by adding one to the right coordinate.

@@

% See https://github.com/jeromekelleher/sc2ts-paper/issues/472
\begin{table}
\caption{Major deletions in Li et al. (2023).}
\label{tab:indels}
\begin{center}
\begin{tabular}{rrcrrrr}
\toprule 
& \multicolumn{4}{c}{Li et al. (2023)} & \multicolumn{2}{c}{Sc2ts ARG} \\
\cmidrule(lr){1-5} \cmidrule(lr){6-7}
Start & Length & Region & Count & Frequency (\%) & Count & Frequency (\%)\\
\midrule
28362 & 9 & N & 3939164 & 43.0525 & 871774 & 35.1216 \\
28271 & 1 & Intergenic & 3853692 & 42.1183 & 1118363 & 45.0561 \\
21765 & 6 & S & 3179887 & 34.7541 & 339681 & 13.6849 \\
11288 & 9 & ORF1ab-nsp6 & 3040576 & 33.2315 & 290408 & 11.6998 \\
28248 & 6 & ORF8 & 3026186 & 33.0742 & 1114669 & 44.9073 \\
22029 & 6 & S & 3010665 & 32.9046 & 1112410 & 44.8163 \\
11283 & 9 & ORF1ab-nsp6 & 2011934 & 21.9891 & 40 & 0.0016 \\
6513 & 3 & ORF1ab-nsp3 & 1984683 & 21.6913 & 340765 & 13.7286 \\
21633 & 9 & S & 1941962 & 21.2244 & 532072 & 21.4359 \\
21987 & 9 & S & 1918094 & 20.9635 & N/A & N/A \\
22194 & 3 & S & 1833050 & 20.0340 & 312706 & 12.5982 \\
21991 & 3 & S & 912944 & 9.9779 & 288831 & 11.6363 \\
686 & 9 & ORF1ab-nsp1 & 115450 & 1.2618 & N/A & N/A \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
